Analyze Stop Data

### 2016-08-08 eda/CensusDataExploration/HeatmapExploration2.Rmd

@@

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
   select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

# %>%
#   mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
#   mutate(Relative.Per.Other = Other/sum(Other)) %>%
#   mutate(Relative.Per.White = White/sum(White)) %>%
#   mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
#   mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%

names(bpop2)[2] <- "Population"
names(bpop2)[3] <- "Hispanic"
names(bpop2)[5] <- "Black"
names(bpop2)

@@

### BPD Stops:
```{r}
# edit bpd stops:
# nblacks <- sum(df$Race == "Black")
# nasians <- sum(df$Race == "Asian")
# nwhite <- sum(df$Race == "White")
# nother <- sum(df$Race == "Other")
# nhisp <- sum(df$Race == "Hispanic")
# totalpop <- sum(berkpop$TotalPop)
# stop.summary <- data.frame(Race = unique(df$Race))

stop.summary <- df %>%  
   group_by(Race) %>%
   summarise(Stopped=n() ) %>%
  mutate(Percent.Stopped = Stopped/nrow(df))

racepops <- c(sum(bpop2$Asian), sum(bpop2$Black), sum(bpop2$Hispanic), sum(bpop2$Other), sum(bpop2$White) )

stop.summary$Race.Population = racepops
stop.summary$Percent.Race.Stopped = stop.summary$Stopped/stop.summary$Race.Population

```{r}
# all stop density:
allbpd_density <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = 2* ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank()) +
  guides(alpha = FALSE)
allbpd_density

# by race:
bpd_race <- berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha =  ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradientn(trans = "log10", colors = p) +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  ) + 
  guides(alpha = FALSE) +
  facet_grid( ~ Race)
bpd_race

### Bar Charts:

```{r}
p <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  facet_grid(~Race) +
  scale_y_continuous(labels=percent) +
  labs(y = "Percent Stopped") +
  ggtitle("Results of Stops by Race")
p
```

```{r}
p1 <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") + ggtitle("Results of Stops") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p1race <- ggplot(df, aes(Enforcement, fill = Enforcement)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Results of Stops by Race") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

```{r}
grid.arrange(p1,p1race,ncol=2)
```

```{r}

p2 <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
+  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent") + ggtitle("Reason for Stops") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p2race <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Reason for Stops") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p3 <- ggplot(df, aes(CarSearch, fill = CarSearch)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Car Search by Race") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )   + guides(fill = FALSE)

p3race <- ggplot(df, aes(Reason, fill = Reason)) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Car Search by Race") +
  facet_grid( ~ Race) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

df$AgeRange = factor(df$AgeRange)
levels(df$AgeRange) <- c("10-19","20-29","30-39","40-49")

p4 <- ggplot(df, aes(AgeRange, fill = AgeRange )) + 
    geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) +  
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(y = "Percent Stopped") +
  ggtitle("Age Range of Car Stops by Race") +
  labs(x = "Age Range") +
  labs(fill = "Age Range")

```{r}

>>>>

### All BPD Stops Density, 2015-2016
```{r}
# a contour plot
berkgg +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
                 size = 2, bins = 5, data = df, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
  theme (panel.grid.major = element_blank (), # remove major grid
         panel.grid.minor = element_blank ()  # remove minor grid
  )+ 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)

### Works, but no google maps:
```{r, eval = F, include = F}
### Read in .shp file only for Census:
shpfile <- str_c(census_tract_folder, "/", census_tract_shp, ".shp")
sh <- readShapePoly(shpfile)
```

```{r, eval = F, include = F}
### Plot the census shape file (polygon outline only)
plot(sh)
```

```{r, eval = F, include = F}
### Modify the shape file to plot the population info:
sh@data$ID <- as.numeric(sh@data$ID)

# Read in the demographic data and merge on Neighbourhood Id
census5_2 <- read.csv(file=census_tract_csv, header = TRUE)
sh2 <- merge(sh, census5_2, by='ID')

# Set the palette
p <- colorRampPalette(c("white", "red"))(128)
palette(p)

# Scale the total population to the palette
popamts <- sh2@data$TotalPop.x

# create the colors:
cols <- (popamts - min(popamts))/diff(range(popamts))*127+1
```

```{r, eval = F, include = F}
### Plot the census shape data with colors for population:
plot(sh, col=cols)
```

### Cleaning the Berkeley population info (do not run):
```{r, eval = F, include = F}
#write.csv(alameda, file = "alameda_population_by_tract.csv")
# alameda = alameda[-1,]
# rownames(alameda) <- 1:nrow(alameda)
# names(alameda)[1] <- "GEO"
# 
# alameda2 <- alameda %>%
#   separate(GEO, c("Block","Group", "Tract", "County", "State"), ", ")
# 
# names(alameda2)[3] <- "NAME10"
# 
# alameda2 <- alameda2 %>%
#   mutate(NAME10 = str_replace(NAME10, "Census Tract ", ""))
# 
# alameda2 <- alameda2 %>%
#   select(-Block,-Group,-County,-State) 
# 
# a3 <- alameda2 %>%
#         group_by(NAME10)%>% 
#         summarise(Total.population = sum(Total.population),
#                   Hispanic.or.Latino = sum(Hispanic.or.Latino),
#                   Total.population..not.Hispanic.or.Latino = sum(Total.population..not.Hispanic.or.Latino),
#                   One.race.total = sum(One.race.total),
#                   White = sum(White),
#                   Black.or.African.American = sum(Black.or.African.American),
#                   American.Indian.and.Alaska.Native = sum(American.Indian.and.Alaska.Native),
#                   Asian = sum(Asian),
#                   Native.Hawaiian.and.Other.Pacific.Islander = sum(Native.Hawaiian.and.Other.Pacific.Islander),
#                   Some.Other.Race = sum(Some.Other.Race),
#                   Two.or.More.Races = sum(Two.or.More.Races))

#census_alameda <- left_join(censusCSV, a3, by = "NAMELSAD10")
```

```{r, eval = F, include = F}
# cleaning the berkeley population data, do not run me:

# census.df <- fortify(censusT, region = 'NAME10')
# pop <- censusCSV
# pop <- pop %>%
#   select(NAME10, TotalPop) %>%
#   mutate(NAME10 = as.character(NAME10))
# 
# names(a3)[1] <- "id"
# names(pop)[1] <- "id"
# 
# pop <- left_join(pop, a3, by = "id")
# census.df <- left_join(census.df, pop, by = "id")
#write.csv2(a3,"alameda_clean_census.csv")
#write.csv2(pop, "berkeley_census_clean.csv")
```

@@

bpop2 <- bpop2 %>%
  mutate(Percent.Black = Black.or.African.American/Total.population) %>%
  mutate(Percent.Other = Other/Total.population) %>%
  mutate(Percent.White = White/Total.population) %>%
  mutate(Percent.Asian = Asian/Total.population) %>%
  mutate(Percent.Hispanic = Hispanic.or.Latino/Total.population) %>%
  mutate(Percent.Berkeley = Total.population/sum(Total.population)) %>%
  select(-TotalPop) %>%
  mutate(Relative.Per.Black = Black.or.African.American/sum(Black.or.African.American)) %>%
  mutate(Relative.Per.Other = Other/sum(Other)) %>%
  mutate(Relative.Per.White = White/sum(White)) %>%
  mutate(Realtive.Per.Asian = Asian/sum(Asian)) %>%
  mutate(Relative.Per.Hispanic = Hispanic.or.Latino/sum(Hispanic.or.Latino)) %>%
  select(-American.Indian.and.Alaska.Native,-Native.Hawaiian.and.Other.Pacific.Islander, -Some.Other.Race)

@@

### 2016-08-08 eda/PaperDrafts/Some thoughts.Rmd

############################## stop data ############################## 
stop_by <- read_csv("cleaned_stop_data.csv")

stop_by$Call.Date.Time <- mdy_hm(stop_by$Call.Date.Time)
stop_by$AgeRange <- as.factor(stop_by$AgeRange)

stop_by <- stop_by %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

stop_by$Race<- stop_by$Race %>% 
  str_replace("A", "Asian") %>% 
  str_replace("B", "Black") %>% 
  str_replace("H", "Hispanic") %>% 
  str_replace("O", "Other") %>% 
  str_replace("W", "White")

stop_by$Gender<- stop_by$Gender %>% 
  str_replace("F", "Female") %>% 
  str_replace("M", "Male")

stop_by$Reason <- stop_by$Reason %>% 
  str_replace("I", "Investigation") %>% 
  str_replace("T", "Traffic") %>% 
  str_replace("R", "Reasonable Suspicion") %>% 
  str_replace("R", "Probation/ Parole") %>% 
  str_replace("W", "Wanted")

stop_by$Enforcement <- stop_by$Enforcement %>% 
  str_replace("A", "Arrest") %>% 
  str_replace("C", "Citation") %>% 
  str_replace("O", "Other") %>% 
  str_replace("W", "Warning")

stop_by$CarSearch <- stop_by$CarSearch %>% 
  str_replace("S", "Search") %>% 
  str_replace("N", "No Search")

@@

@@

```{r}
p1 <- ggplot(stop_by)+
  geom_bar(aes(x=Race),
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Stop data by race",
       x = "Race",
       y = "Counts")

p2 <- ggplot(race)+
  geom_bar(aes(x=Description, 
               y=Counts),
           stat = "identity", 
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Demographic race data",
       x = "Race",
       y = "Counts")

multiplot(p1, p2,cols=2)

>>>>

@@

mergedf$Call.Date.Time <- mdy_hm(mergedf$Call.Date.Time)
mergedf$AgeRange <- as.factor(mergedf$AgeRange)

mergedf <- mergedf %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

mergedf$Gender<- mergedf$Gender %>% 
  str_replace("F", "Female") %>% 
  str_replace("M", "Male")

mergedf$AgeRange<- mergedf$AgeRange %>% 
  str_replace("1", "0-18") %>% 
  str_replace("2", "18-29") %>% 
  str_replace("3", "30-39") %>%  
  str_replace("4", "40+")

@@

```{r, fig.height = 5, fig.width = 10, echo = FALSE}

m <- stop_by %>% 
  group_by(Race) %>% 
  tally()
names(m) <- c("Race", "stop")
names(race) <- c("Race", "census")

pop <- left_join(race,m,by = "Race")
pop <- pop %>% 
  mutate(percentage=(stop/census)) %>% 
pop$percentage<- round(pop$percentage, digits = 2)

ggplot(pop)+
  geom_bar(aes(x=Race,
               y=percentage), 
           stat = "identity",
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40b8d0"))+
  geom_text(aes(x = Race, 
                y = percentage, 
                label = paste0(percentage,"%")), 
            size=5,
            colour= "#000000")+
  scale_y_continuous(labels=percent, limits = c(0,1)) +
  labs(title = "The percentage of being stopped by police (%)",
       x = "Race",
       y = "Percentage") +
  theme(axis.line = element_line(size=1, colour = "black"), 
        panel.grid.major = element_line(colour = "#d3d3d3"), panel.grid.minor = element_blank(), 
        panel.border = element_blank(), panel.background = element_blank()) + 
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"), 
        text=element_text(family="Tahoma"),
        axis.text.x=element_text(colour="black", size = 10), 
        axis.text.y=element_text(colour="black", size = 10))

```{r}
race %>%
  ggvis(x = ~Description, y = ~Counts) %>%
  layer_bars(fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")) %>%
  add_tooltip(function(race) race$Counts) 
  labs(title = "Counts of people recorded in every day in a week")

### 2016-08-08 eda/PaperDrafts/draft.Rmd

---
title: "BDP draft paper"
author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
date: "2 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(readr)
library(plyr)
library(dplyr)
library(ggvis)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
library(grid)
library(gridExtra)
```

@@

* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
* [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)

@@

```{r, echo=FALSE}
# Read data

@@

p1 <- ggplot(stop_by)+
  geom_bar(aes(x=Race),
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Stop data by race",
       x = "Race",
       y = "Counts")

p2 <- ggplot(race)+
  geom_bar(aes(x=Description, 
               y=Counts),
           stat = "identity", 
           fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF"))+
  labs(title = "Demographic race data",
       x = "Race",
       y = "Counts")

@@

@@

@@

>>>>

---
title: "Berkeley Police Data"
author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(readr)
library(plyr)
library(dplyr)
library(ggvis)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
library(grid)
library(gridExtra)
```

@@

* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
* [Berkeley Census Data](https://data.cityofberkeley.info/Demographics/Census-Data-2000-And-2010/bnkq-f2ge)
*[Census 2010 Population and race data by county by census tract](http://www.bayareacensus.ca.gov/small/small.htm)

@@

```{r, echo=FALSE, include = FALSE, message=FALSE, warning = FALSE}
# Read data

@@

p1 <- ggplot(mergedf)+
  geom_bar(aes(x=Race),
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  labs(title = "Stop data by race",
       x = "Race",
       y = "Counts")

p2 <- ggplot(race)+
  geom_bar(aes(x=Race, 
               y=census),
           stat = "identity", 
           fill = c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0"))+
  labs(title = "Demographic race data",
       x = "Race",
       y = "Counts")

@@

c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")

@@

c("#FFCC00","#E1B378","#5F9EA0","#66FFCC","#40B8D0")

@@

```{r}
race %>%
  ggvis(x = ~Description, y = ~Counts) %>%
  layer_bars(fill = c("#FFCC00","#000000","#330099","#66FFCC","#FFFFFF")) %>%
  add_tooltip(function(race) race$Counts) 
  labs(title = "Counts of people recorded in every day in a week")

### 2016-08-08 eda/StopDataExploration/AnalysizeStop.Rmd

>>>>

```{r}
stop_reason <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
   geom_bar(aes(y = ..count../sum(..count..))) + 
   labs(y = "Density") +
   scale_fill_brewer(palette = "Set1") +
   facet_grid(AgeRange~Gender)
stop_reason
```

```{r}
stop_enforcement <- ggplot(stop_by, aes(x = Reason, fill = Enforcement)) + 
  geom_bar(aes(y = ..count../sum(..count..))) + 
  labs(y = "Density") +
  facet_wrap(~Race)
stop_enforcement
```
